Endogenous Politics and \\ the Design of Trade Institutions

### 2017-01-17 draft.tex
integrate more sign-posting to MRC2007

We would like a framework in which the government's payoff is not constant. An appealing candidate is \Textcite{mrc2007}, which is in the spirit of \Textcite{gh94} but uses a bargaining framework. In my setup, the lobby is not active during the trade agreement negotiation; in order to to fit into the framework of \Textcite{mrc2007}, we can assume that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement.

Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights.

In \Textcite{lt}'s setup, the government, with objective function $G = \Psi(e) + aW(\tau,\tau^*)$, decides whether to commit in an international agreement and then Nash bargains with the lobby over $e$ and the tariff.\footnote{I have changed notation to match that which I use here and restricted attention to their Section 3.3. with the tariff only; their main results address the choice between tariffs and non-tariff barriers.} They show that the government will use the trade agreement to commit vis-\`{a}-vis the lobby as long as it doesn't have all the bargaining power and there are decreasing returns to lobbying. Here, the government does not have an interest in maximizing joint surplus because it is unable to appropriate all the gains.\footnote{In the working paper version of \Textcite{mrc2007}, a similar result is demonstrated for the case of immobile capital when the government has all the bargaining power.}

@@

With the introduction of endogenous political pressure, one must take a stand on whether lobbies influence the formation of the trade agreement or are only active once the trade agreement is in place. Since a central goal here is to determine the impact of relaxing the assumption that political pressure is exogenously determined, I assume the latter to make the comparison most direct to this literature in which by definition there can be no endogenous political pressure at the trade agreement formation stage. The former, which is undoubtedly an important possibility in many institutional settings, is left as an important extension.\footnote{Note that the endogenous trade policy literature that involves trade agreements (e.g. \Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure while answering a different set of questions. Results for the case of ex-ante lobbying are available from the author upon request.}

@@

Thus we see that the governments can use a tariff cap to manipulate lobbying incentives so that political effort is neither too low or too high. Note that this is related to a point made by \Textcite{mrc2007}; in their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.

>>>>

We would like a framework in which the government's payoff is not constant. An appealing candidate is \Textcite{mrc2007}, which is in the spirit of \Textcite{gh94} but uses a bargaining framework. In my setup, the lobby is not active during the trade agreement negotiation; in order to to fit into the framework of \Textcite{mrc2007}, we can assume that capital is perfectly mobile in the long run so that it is not worthwhile for the lobby to expend resources to influence the negotiation of the trade agreement. Of course, I consider only the case of perfectly immobile capital in the short run, whereas \Textcite{mrc2007} examines all possible levels of capital mobility.

Although \Textcite{mrc2007} also assumes the constant returns form and that either the lobby or the government has all the bargaining power, \Textcite{lt} extend this framework to one with decreasing returns to lobbying and general bargaining weights in an environment with no ex-ante lobbying.

In \Textcite{lt}'s setup, the government, with objective function $G = \Psi(e) + aW(\tau,\tau^*)$, decides whether to commit in an international agreement and then Nash bargains with the lobby over $e$ and the tariff.\footnote{I have changed notation to match that which I use here and restricted attention to their Section 3.3. with the tariff only; their main results address the choice between tariffs and non-tariff barriers.} They show that the government will use the trade agreement to commit vis-\`{a}-vis the lobby as long as it doesn't have all the bargaining power and there are decreasing returns to lobbying. Here, the government does not have an interest in maximizing joint surplus because it is unable to appropriate all the gains.\footnote{In the working paper version of \Textcite{mrc2007}, a similar result is demonstrated for the case of immobile capital when the government has all the bargaining power and there is no ex-ante lobbying.}

@@

With the introduction of endogenous political pressure, one must take a stand on whether lobbies influence the formation of the trade agreement or are only active once the trade agreement is in place. Since a central goal here is to determine the impact of relaxing the assumption that political pressure is exogenously determined, I assume the latter to make the comparison most direct to this literature in which by definition there can be no endogenous political pressure at the trade agreement formation stage. The former, which is undoubtedly an important possibility in many institutional settings, is left as an important extension.\footnote{Note that the endogenous trade policy literature that involves trade agreements (e.g. \Textcite{gh95}, \Textcite{mrc2007}) explicitly assumes that trade agreements are formed in the face of political pressure. Results for the case of ex-ante lobbying are available from the author upon request.}

@@

Thus we see that the governments can use a tariff cap to manipulate lobbying incentives so that political effort is neither too low or too high. The fact that the applied tariffs are equal to the negotiated binding is reminiscent of \Textcite{mrc2007}. Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored. At the same time, in both models the government typically wants to use the trade agreement tariff cap to reduce lobbying below the lobby's optimum. In the model of this paper, it is because the government gets disutility from very high lobbying expenditure; in \Textcite{mrc2007}, it is because high tariff rates encourage inefficient over-investment in the import-competing industry.

%In their model, tariff caps and exact tariff commitments are equivalent when factors are perfectly immobile as in this model.

### 2017-01-17 to-do.tex
integrate more sign-posting to MRC2007

\section{BS2005 vs. MRC2007}
	\begin{itemize}
		\item I like the way you use Bagwell and Staiger (2005) as your benchmark.  But I also wondered whether Maggi and Rodriguez-Clare (2007, henceforth MRC) also offered a useful benchmark.  I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.  In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.  But your mechanisms seem quite different.  They have capital mobility whereas your factors are specific to each sector.  If capital is highly immobile they have a hold-up problem, which creates an incentive for lobbies to try to prevent too much liberalization under the agreement.  What drives this feature in your model?  Here I think it would be useful to explain carefully the sequence of events in your set-up more carefully.  What is the relationship between lobby effort level and tariff policy?  Is effort fixed first, after which the government sets its tariff?  If so it seems to me that would drive a kind of hold-up because lobbying effort is sunk.  It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC. 
		\item DGH / MRC analogy from JIE$\_$revision
			\begin{itemize}
				\item After Prop1 is a good place to start to put references to MRC 2007

				\item The fact that the applied tariffs are equal to the negotiated binding is reminiscent of Maggi and Rodriguez-Clare 2007. Exactly the same dynamic is at play here: specifying trade agreement tariffs as caps instead of strong bindings keeps the lobby active during periods where the trade agreement is honored. In Appendix B.3, I analyze the model with strong bindings and show that the results would be altered in magnitude but not in spirit by assuming that the trade agreement tariffs are strong bindings instead of tariff caps. The only change to the model is that under strong bindings there would be zero lobbying effort during a trade agreement phase as the lobby would not need to exert effort to bid protection levels up to the trade agreement tariff. There would still be no disputes in equilibrium, but the lobbying constraint is easier to satisfy with strong bindings because the gap between trade war and trade agreement profits shrinks when the lobby stops exerting effort to receive the trade agreement tariff. Because lower trade agreement tariffs can be sustained under strong bindings, executive and legislators with small political economy weights prefer a strong-binding agreement and legislators with high political economy weights prefer a weak-binding agreement. A tariff cap makes the lobby's self-enforcement constraint harder to satisfy and thus requires a higher trade agreement tariff for self-enforcement. From the ex-ante point of view, strong bindings are therefore preferable. Tariff caps could be viewed as a way to commit to setting higher trade agreement tariffs and therefore as a mechanism for ensuring that rents are distributed to protectionists ex-post.
			\end{itemize}
		\item MRC07
			\begin{itemize}
				\item Need case of fixed capital ($z=0$)
				\item NB (p. 1391): optimal trade agreement if lobbies have all the bargaining power and no ex-ante lobbying is free trade
					\begin{itemize}
						\item Page 26-28 of working paper has government with all the bargaining power (and no capital mobility, but remember that the capital owners choose where to invest in the first period). Here, could be free trade but could be interior to get some rents.
					\end{itemize}		
			\end{itemize}
	\end{itemize}

@@

\newpage
\begin{center}
\textbf{\large{For Later}}
\end{center}
\begin{enumerate}
	\item Add export lobby. 
		\begin{itemize}
			\item Extension or appendix b/c want to make comparison to existing literature.
		\end{itemize}
\end{enumerate}
\end{document}

>>>>

@@

\newpage
\begin{center}
\textbf{\large{For Later}}
\end{center}
\begin{enumerate}
	\item Add export lobby. 
		\begin{itemize}
			\item Extension or appendix b/c want to make comparison to existing literature.
		\end{itemize}
\end{enumerate}

\end{document}
